---
title: "Shiny Example - Visualizing WDI Data"
author: Thomas Brambor
job: QMSS Data Visualization
widgets: []
mode: selfcontained
always_allow_html: yes
runtime: shiny

output:
  ioslides_presentation:
    incremental: false
    smaller: false
  github_document:
    toc: true
    toc_depth: 1
---

```{r packages, message = FALSE, warning = FALSE, echo=FALSE}
# Load packages.
packages <- c("devtools","knitr","leaflet","shiny","raster","rgdal","RColorBrewer","dplyr","readr")

packages <- lapply(packages, FUN = function(x) {
  if(!require(x, character.only = TRUE)) {
    install.packages(x)
    library(x, character.only = TRUE)
  }
})

```


```{r Setup, include=FALSE, results='hide', warning=FALSE}

# A Prefix nulling hook.

# Make sure to keep the default for normal processing.
default_output_hook <- knitr::knit_hooks$get("output")

# Output hooks handle normal R console output.
knitr::knit_hooks$set( output = function(x, options) {

  comment <- knitr::opts_current$get("comment")
  if( is.na(comment) ) comment <- ""
  can_null <- grepl( paste0( comment, "\\s*\\[\\d?\\]" ),
                     x, perl = TRUE)
  do_null <- isTRUE( knitr::opts_current$get("null_prefix") )
  if( can_null && do_null ) {
    # By default R print output aligns at the right brace.
    align_index <- regexpr( "\\]", x )[1] - 1
    # Two cases: start or newline
    re <- paste0( "^.{", align_index, "}\\]")
    rep <- comment
    x <- gsub( re, rep,  x )
    re <- paste0( "\\\n.{", align_index, "}\\]")
    rep <- paste0( "\n", comment )
    x <- gsub( re, rep,  x )
  }

  default_output_hook( x, options )

})

knitr::opts_template$set("kill_prefix"=list(comment=NA, null_prefix=TRUE))


## Normal Setup From Here
library(knitr)

opts_chunk$set(fig.path="images/",
               cache.path="cache/",
            #   dev=c("png","pdf"),
            #  fig.width=5,
            #  fig.height=4.5,
            #  dpi=300,
            #  fig.show="hold",
            #  fig.lp="fig:",
               cache=TRUE,
            #  par=TRUE,
               echo=TRUE,
               message=FALSE,
               warning=FALSE)
```


# Starting with the template

```{r, echo=FALSE, cache=FALSE, eval=TRUE}
library(shiny)
ui <- fluidPage()
server <- function(input, output) {}
shinyApp(ui = ui, server = server)
```

## Add some layout

```{r, echo=FALSE, cache=FALSE, eval=TRUE}
library(shiny)

ui <- fluidPage(
  titlePanel("Visualizing the World Development Indicators"),
  
  sidebarLayout(sidebarPanel("sidebar panel"),
                mainPanel("main panel"))
)

server <- function(input, output) {
}

shinyApp(ui = ui, server = server)
```

## Define input lists

```{r, echo=FALSE, cache=FALSE, eval=TRUE}
# Content of our SETUP.R file
library(WDI)
library(readxl)

# Set up the data
# Generate a lookup table
# source: databank.worldbank.org/data/download/site-content/WDI_CETS.xls

xl <- "WDI_CETS.xls"

lookup <- read_excel(xl)
names(lookup) <- c("a", "b", "c", "d", "e", "f")

# Functions to lookup variable names / descriptions
lookup_list  <- setNames(as.list(lookup$b), lookup$a)
lookup_list2 <- setNames(as.list(lookup$a), lookup$b)

# Lookup countries
cc <- read_csv("country_codes.csv", na="", 
               col_types = list(col_factor(), col_factor()))

lookup_list_countries  <- setNames(as.list(cc$iso_2c), cc$cname)
lookup_list_countryabb <- setNames(as.list(cc$cname), cc$iso_2c)
``` 

## Define input variables

```{r, echo=FALSE, cache=FALSE, eval=TRUE}
## The source file creates the lookup lists
source('setup.R')

ui <- fluidPage(
  titlePanel("Visualizing World Development Indicators"),
  
  sidebarLayout(
    sidebarPanel(
      title = "Make selections",
      selectInput(
        'indicator',
        label = 'World Development Indicator',
        c("Variable 1", "Variable 2", "Variable3"),
        selected = 'Variable 1'
      ),
      selectInput(
        "ctry",
        label = "Country",
        c("US", "CA", "FR"),
        selected = "US"
      )
    ),
    mainPanel(
      "Visualization",
    )
  )
)

server <- function(input, output) {
}

shinyApp(ui = ui, server = server)
```

## Named List instead

```{r, echo=FALSE, cache=FALSE, eval=TRUE}
source('setup.R')

ui <- fluidPage(
  titlePanel("Visualizing World Development Indicators"),
  
  sidebarLayout(
    sidebarPanel(
      title = "Make selections",
      selectInput(
        'indicator',
        'World Development Indicator',
        lookup_list2,
        selected = 'Fertility rate, total (births per woman)'
      ),
      selectInput(
        "ctry",
        label = "Country",
        c(
          "United States" = "US",
          "Canada" = "CA",
          "France" = "FR"
        ),
        selected = "US"
      )
    ),
    mainPanel(h2(h1("Visualization")))
  )
)

server <- function(input, output) {
}

shinyApp(ui = ui, server = server)
```

## Adding some simple output for debugging

```{r, echo=FALSE, cache=FALSE, eval=TRUE}
source('setup.R')

ui <- fluidPage(
  titlePanel("Visualizing World Development Indicators"),
  
  sidebarLayout(
    sidebarPanel(
      title = "Make selections",
      selectInput(
        'indicator',
        'World Development Indicator',
        lookup_list2,
        selected = 'Fertility rate, total (births per woman)'
      ),
      selectInput(
        "ctry",
        label = "Country",
        c(
          "United States" = "US",
          "Canada" = "CA",
          "France" = "FR"
        ),
        selected = "US"
      )
    ),
    mainPanel(h2(h1("Visualization")),
              textOutput("selected_indicator"),
              textOutput("selected_country"))
  )
)

server <- function(input, output) {
  output$selected_indicator <- renderText({ 
    paste("You have selected the variable", input$indicator)
  })
  output$selected_country <- renderText({ 
    paste("You have selected the country", input$ctry)
  })
}

shinyApp(ui = ui, server = server)
```


## Could group the input list as well

```{r, echo=FALSE, cache=FALSE, eval=TRUE}
source('setup.R')

ui <- fluidPage(
  titlePanel("Visualizing World Development Indicators"),
  
  sidebarLayout(
    sidebarPanel(
      title = "Make selections",
      selectInput(
        'indicator',
        'World Development Indicator',
        lookup_list2,
        selected = 'Fertility rate, total (births per woman)'
      ),
      selectInput(
        "ctry",
        label = "Country",
        list(
          `North America` = c("United States" = "US",
                              "Canada" = "CA"),
          
          `Europe` = c("France" = "FR",
                       "Germany" = "DE")
        ),
        selected = "US"
      )
    ),
    mainPanel(h1("Visualization"))
  )
)

server <- function(input, output) {
}

shinyApp(ui = ui, server = server)
```


## Define input variables: now country as lookup as well

```{r, echo=FALSE, cache=FALSE, eval=TRUE}
source('setup.R')

ui <- fluidPage(
  titlePanel("Visualizing World Development Indicators"),
  
  sidebarLayout(
    sidebarPanel(
      title = "Make selections",
      selectInput(
        'indicator',
        'World Development Indicator',
        lookup_list2,
        selected = 'Fertility rate, total (births per woman)'
      ),
      selectInput(
        "ctry",
        label = "Country",
        lookup_list_countries,
        selected = "US",
      )
    ),
    mainPanel(
      h2(h1("Visualization")),
      textOutput("selected_indicator"),
      textOutput("selected_country")
    )
  )
)

server <- function(input, output) {
  output$selected_indicator <- renderText({
    paste("You have selected the variable", input$indicator)
  })
  output$selected_country <- renderText({
    paste("You have selected the country", input$ctry)
  })
}

shinyApp(ui = ui, server = server)
```

## Define output in UI

```{r, echo=FALSE, cache=FALSE, eval=TRUE}
source('setup.R')

ui <- fluidPage(
  titlePanel("Visualizing World Development Indicators"),
  
  sidebarLayout(
    sidebarPanel(
      title = "Make selections",
      selectInput(
        'indicator',
        'World Development Indicator',
        lookup_list2,
        selected = 'Fertility rate, total (births per woman)'
      ),
      selectInput(
        "ctry",
        label = "Country",
        lookup_list_countries,
        selected = "US"
      )
    ),
    
    mainPanel(h2(h1("Visualization")),
              dygraphOutput("dchart", width = "100%"))
  )
)


server <- function(input, output) {
}

shinyApp(ui = ui, server = server)
```

## Define inputs in server: Data

This is how the `WDI` package works:

```{r, echo=TRUE, cache=FALSE, eval=FALSE}
df <- WDI(
  country = "US",
  indicator = "AG.LND.ARBL.HA",
  start = 1980,
  end = 2016,
  extra = FALSE
)

# To transform the data
df1 <- df %>%
  dplyr::select("country", "year", "AG.LND.ARBL.HA") %>%
  tidyr::spread(key = "country", value = "AG.LND.ARBL.HA") %>%
  dplyr::mutate(date = as.Date(as.character(year),
                               format = "%Y")) %>%
  dplyr::select(-year) 
```

## Define inputs and outputs in server: Data and Graph

```{r, echo=FALSE, cache=FALSE, eval=TRUE}
source('setup.R')

ui <- fluidPage(
  titlePanel("Visualizing World Development Indicators"),
  
  sidebarLayout(
    sidebarPanel(
      title = "Make selections",
      selectInput(
        'indicator',
        'World Development Indicator',
        lookup_list2,
        selected = 'Fertility rate, total (births per woman)'
      ),
      selectInput(
        "ctry",
        label = "Country",
        lookup_list_countries,
        selected = "US"
      )
    ),
    mainPanel(h2(h1("Visualization")),
              dygraphOutput("dchart", width = "100%"))
  )
)

server <- function(input, output) {
  output$dchart <- renderDygraph({
    # Which variable to get?
    ind <- input$indicator
    # Get the data
    df <- WDI(
      country = input$ctry,
      indicator = ind,
      start = 1980,
      end = 2016,
      extra = FALSE
    )
    # Transform the data
    df1 <- df %>%
      dplyr::select("country", "year", ind) %>%
      tidyr::spread(data = ., key = "country", value = ind) %>%
      dplyr::mutate(date = as.Date(as.character(year), format = "%Y")) %>%
      dplyr::select(-year)
    
    xtdata <- xts(df1, order.by = df1$date)
    xtdata$date <- NULL
    
    dygraph(xtdata) %>%
      dyOptions(colors = c("black", "black"))
  })
  
}

shinyApp(ui = ui, server = server)
```

## Add range selector

```{r, echo=FALSE, cache=FALSE, eval=TRUE}
source('setup.R')

ui <- fluidPage(
  titlePanel("Visualizing World Development Indicators"),
  
  sidebarLayout(
    sidebarPanel(
      title = "Make selections",
      selectInput(
        'indicator',
        'World Development Indicator',
        lookup_list2,
        selected = 'Fertility rate, total (births per woman)'
      ),
      selectInput(
        "ctry",
        label = "Country",
        lookup_list_countries,
        selected = "US"
      )
    ),
    mainPanel(h2(h1("Visualization")),
              dygraphOutput("dchart", width = "100%"))
  )
)

server <- function(input, output) {
  output$dchart <- renderDygraph({
    # Which variable to get?
    ind <- input$indicator
    # Get the data
    df <- WDI(
      country = input$ctry,
      indicator = ind,
      start = 1980,
      end = 2016,
      extra = FALSE
    )
    # Transform the data
    df1 <- df %>%
      dplyr::select_("country", "year", ind) %>%
      tidyr::spread(data = ., key = "country", value = ind) %>%
      dplyr::mutate(date = as.Date(as.character(year), format = "%Y")) %>%
      dplyr::select(-year)
    
    xtdata <- xts(df1, order.by = df1$date)
    xtdata$date <- NULL
    
    # Name of country
    countryname    <- lookup_list_countryabb[[input$ctry]]
    ind_desc       <- lookup_list[["AG.LND.ARBL.HA"]]
    dygraph(xtdata,
            main = paste(ind_desc, "in", countryname),
            xlab = "Year") %>%
      dyOptions(colors = c("black", "black")) %>%
      dyRangeSelector(height = 20)
  })
  
}

shinyApp(ui = ui, server = server)
```

## Add highlighting for mouse over

```{r, echo=FALSE, cache=FALSE, eval=TRUE}
source('setup.R')

ui <- fluidPage(
  titlePanel("Visualizing World Development Indicators"),
  
  sidebarLayout(
    sidebarPanel(title="Make selections",
               selectInput('indicator', 
                           'World Development Indicator', 
                           lookup_list2, 
                           selected = 'Fertility rate, total (births per woman)'), 
               selectInput("ctry", label = "Country", 
                           lookup_list_countries, 
                           selected = "US") 
                 ),
    mainPanel(h1("Visualization"),
              dygraphOutput("dchart", width = "100%")
             )
  )
)

server <- function(input, output) {
  
    output$dchart <- renderDygraph({ 
      # Which variable to get?  
      ind <- input$indicator
      # Get the data
      df <- WDI(country = input$ctry, indicator = ind, 
               start = 1980, end = 2016, extra = FALSE)
      # Transform the data
      df1 <- df %>%
        dplyr::select_("country", "year", ind) %>%
        tidyr::spread(data = ., key = "country", value = ind) %>%
        dplyr::mutate(date = as.Date(as.character(year), format = "%Y")) %>%
        dplyr::select(-year) 
    
    xtdata <- xts(df1, order.by = df1$date) 
    xtdata$date <- NULL
    
    # Name of country
    countryname    <- lookup_list_countryabb[[input$ctry]]
    ind_desc       <- lookup_list[["AG.LND.ARBL.HA"]]
    dygraph(xtdata,
            main = paste(ind_desc,"in",countryname),
            xlab = "Year") %>%
      dyOptions(colors = c("black", "black")) %>%
      dyRangeSelector(height = 20) %>%
      dyHighlight(highlightCircleSize = 5, 
              highlightSeriesBackgroundAlpha = 0.2,
              hideOnMouseOut = FALSE)
    })
  
}

shinyApp(ui = ui, server = server)
```

## Adding styling with CSS

- we can add styling to one particular graph only by attaching CSS file specifically for the graph with 

```
dygraph(...) %>%
  dyCSS("dygraph.css")
```

Here are the relevant CSS tags: http://dygraphs.com/css.html

## Adding styling with CSS

A better option is an **overall CSS stylesheet** change how the Shiny app looks like. Just add it to the `head()` tag in the UI:


```{r, echo=FALSE, cache=FALSE, eval=TRUE}
source('setup.R')

ui <- fluidPage(
  tags$head(includeCSS("styles.css")),
  titlePanel("Visualizing World Development Indicators"),
  
  sidebarLayout(
    sidebarPanel(title="Make selections",
               selectInput('indicator', 
                           'World Development Indicator', 
                           lookup_list2, 
                           selected = 'Fertility rate, total (births per woman)'), 
               selectInput("ctry", label = "Country", 
                           lookup_list_countries, 
                           selected = "US") 
                 ),
    mainPanel(h2(h1("Visualization")),
              dygraphOutput("dchart", width = "100%")
             )
  )
)

server <- function(input, output) {
  
    output$dchart <- renderDygraph({ 
      # Which variable to get?  
      ind <- input$indicator
      # Get the data
      df <- WDI(country = input$ctry, indicator = ind, 
               start = 1980, end = 2016, extra = FALSE)
      # Transform the data
      df1 <- df %>%
        dplyr::select_("country", "year", ind) %>%
        tidyr::spread(data = ., key = "country", value = ind) %>%
        dplyr::mutate(date = as.Date(as.character(year), format = "%Y")) %>%
        dplyr::select(-year) 
    
    xtdata <- xts(df1, order.by = df1$date) 
    xtdata$date <- NULL
    
    # Name of country
    countryname    <- lookup_list_countryabb[[input$ctry]]
    ind_desc       <- lookup_list[["AG.LND.ARBL.HA"]]
    dygraph(xtdata,
            main = paste(ind_desc,"in",countryname),
            xlab = "Year") %>%
      dyOptions(colors = c("black", "black")) %>%
      dyRangeSelector(height = 20) %>%
      dyHighlight(highlightCircleSize = 5, 
              highlightSeriesBackgroundAlpha = 0.2,
              hideOnMouseOut = FALSE)
    })
  
}

shinyApp(ui = ui, server = server)
```

# Back to lecture and later part II






